```{css, echo = FALSE, include = knitr::is_html_output()}
.caption {
  text-align: justify;
  font-size: 0.8em;
  color: #555;
}
```

```{r result-setup, include = FALSE}
library(yaml)
config <- yaml.load_file("config.yml")
fig_dir <- file.path(getwd(), config$fig_dir)

knitr::opts_chunk$set(echo = FALSE)

add_figure <- function(path) {
  if (file.exists(path)) {
    return(knitr::include_graphics(path, auto_pdf = TRUE, rel_path = FALSE))
  }
  warning("File not found: ", path)
  plot(1)
}
```

# Results

```{r team-trends, fig.cap = cap, out.width = "100%", fig.align = "center"}

cap <- "Evolution of Teaching Teams by Gender Composition and Size. Trends in course gender composition from 2000 to 2019 reveal a steady increase in women's participation (both in solo-taught and two-instructor courses), while the proportion of mixed-gender teams has remained relatively stable over time. (A) Percentage of solo-taught courses by instructor gender over time. (B) Percentage of two-instructor courses by gender composition over time. N = 5.1 million courses. Shaded area represent 95\\% confidence interval of the proportion per academic year." # nolint

add_figure(file.path(fig_dir, "evolution-1.png"))
```

Figure \@ref(fig:team-trends) illustrates gender composition and team size trends in university courses for single- and two-instructor courses (2000-2019). Panel A reveals a decline in male-only single-instructor courses from around 60% in 2000 to 46% in 2019, alongside an increase in female-only single-instructor courses from roughly 25% to 36%.[^women-academia] Panel B shows male-only two-instructor courses also decreased, from about 8% in the early 2000s to 6% in 2020, while female-only and mixed-gender two-instructor courses remained relatively stable, each accounting for approximately 4% and 3%, respectively. These patterns indicate a consistent shift toward greater gender diversity in course instruction over time, with mixed-gender teaching teams (3%) consistently less common than same-gender teams (10%).

[^women-academia]: This overall trend aligns with the increase in women holding academic positions worldwide. In 2000, women constituted approximately 35% of academic staff worldwide. By 2022, this figure had risen to about 44%, according to data from the World Bank; The data are available at: https://data.worldbank.org/indicator/SE.TER.TCHR.FE.ZS

```{r montecarlo, fig.cap = cap, out.width = "100%", fig.align = "center", echo = FALSE}
cap <- "Comparison of gender composition between actual courses and courses simulated with Monte Carlo. Simulations ensure gender neutrality while keeping constant the institutional constraints (i.e., number of teams by size, institution, academic year, and 69 academic fields). Simulated mixed-gender teams consistently exceed observed proportions, highlighting persistent barriers to forming gender-diverse teaching teams." # nolint

add_figure(file.path(fig_dir, "montecarlo-1.png"))
```

To assess whether gender imbalances in co-teaching teams persist after accounting for institutional constraints across fields and institutions, Figure \@ref(fig:montecarlo) compares observed teaching team compositions (2000–2019) with those generated by a gender-neutral Monte Carlo simulation. This model randomly shuffles instructors while preserving the overall distribution of course loads per year by institution and academic field but ignoring gender-based preferences or constraints (see Section \@ref(sec:methods)).

The results reveal that actual data systematically deviate from simulated expectations: same-gender teams, such as female-female (FF) and male-male (MM) courses, occur *more* freequently in actual data than in simulations by 66% and 20%, respectively, in 2019. By contrast, mixed-gender collaborations (MF/FM) are consistently *underrepresented* in the actual data (3%) compared to simulations (6%). These findings suggest that significant social or institutional dynamics---beyond chance---reinforce same-gender pairings, especially among women. Such dynamics may include gender-based homophily, mentoring networks, or within-department barriers or assignment practices.

```{r montecarlo-by-field, fig.cap = cap, out.width = "100%", fig.align = "center"}
cap <- "This figure disaggregates the comparison of proportions between actual (green) and simulated (orange) mixed-gender (MF/FM) courses by academic field. The shaded bar indicates 95% confidence level of the proportion." # nolint

add_figure(file.path(fig_dir, "montecarlo-by-field-1.png"))
```

Figure \@ref(fig:montecarlo-by-field) disaggregates the comparison between actual and simulated mixed-gender (MF/FM) courses by academic field. Nearly all fields show clear underrepresentation of mixed-gender teams (relative to expectations) by two to four times. For instance, in Medicine (Health and Welfare), the actual share of mixed-gender collaborations is 4.8% versus an expected 9.1%, indicating a large imbalance. Similar gaps are evident in Law (3.8% actual vs 6.4% simulated) and Linguistic (4.3% vs 6.7%), underscoring the pervasiveness of barriers to mixed-gender teams across different academic fields. Remarkably, we see no differences in these patterns between male- vs female-dominated fields (e.g., Engineering or Accounting vs Nursing or Chemistry), suggesting that the underrepresentation of mixed-gender teams may be a structural feature rather than one driven by field-specific gender imbalances.

```{r montecarlo-by-cntry, fig.cap = cap}
cap <- "This figure disaggregates the comparison of proportions between actual (green) and simulated (orange) mixed-gender courses (MF/FM) by country.  Simulated mixed-gender teams consistently exceed observed proportions in all countries. EU countries are: Austria, Denmark, France, Germany, Ireland, Italy, Netherlands, Poland, Portugal, Spain and Sweden." # nolint

add_figure(file.path(fig_dir, "sim-by-cntry-1.png"))
```

Figure \@ref(fig:montecarlo-by-cntry) illustrates that the proportion of mixed-gender teams in the simulations consistently exceeds the observed proportions across various geographic regions, despite notable variation in the magnitude of these differences. In 2019, for example, mixed-gender teams are relatively rare in Great Britain (1%) compared to Canada (4%). However, in the simulations, the proportion of mixed-gender teams in both countries increases substantially---1% vs 3% and 4% vs 5%, respectively. In 2019, the largest discrepancy between simulated and observed proportions is found in EU countries from 5% to 12%. Overall, these patterns underscore the robustness of the findings across diverse geographic contexts.

## The Impact on Course Materials

To examine how teaching team size and gender composition are associated with important aspects of course materials, we analyse three key metrics: the percentile rank of the interdisciplinarity score per year ("interdisciplinarity"), the average publication year of the readings ("age of readings"), and the share of female authors cited in the assigned readings ("share of female authors"). As discussed before, we use linear mixed-effect regressions to study these relationships while accounting for unobserved differences across academic fields, institutions, time, and other relevant controls.

### Interdisciplinarity

```{r interdisciplinarity, fig.cap = cap, out.width = "100%", fig.align = "center"}
cap <- "This figure illustrates the yearly difference in interdisciplinarity relative to courses taught by a single male instructor (baseline = 0) for each team configuration. The values are regression coefficients obtained separately for each academic year, with controls for country and field. Interdisciplinarity is expressed as the percentile rank of each course’s interdisciplinarity score within its cohort; positive values therefore indicate a greater interdisciplinarity. The shaded area represents 95% confidence intervals." # nolint

add_figure(file.path(fig_dir, "plot-interdisciplinarity-1.png"))
```

Figure \@ref(fig:interdisciplinarity) illustrates the differences in interdisciplinarity across team configurations. Our results indicate that there are no systematical differences in interdisciplinarity between female and male instructors when they are teaching alone: in recent years women tend to be more interdisciplinary, but this effect is not consistent over time.   Conversely, mixed-gender teams tend to exceed in interdisciplinary all-male teams both individual and with two male instructors. This finding is consistent over time. At the same time, we find no consistent evidence of a difference between mixed gender and teams with two female instructors.  These results point to a consistent association between gender diversity and interdisciplinarity.

<!-- Disaggregating the analysis by academic field, with separate regressions by field accounting for year and country controls, shows that the impact of gender diversity is not homogeneous across fields (Figure \@ref(fig:interdisciplinarity-by-field)). Specifically, it appears to have a larger effect in specific fields, especially in Arts and Humanities (between 10%  in 2005 and 5% in 2019) and Social Sciences, Journalism and Information (between 8% and 4% during the same period). -->

### Citing Women Authors

```{r female-authors, fig.cap = cap, out.width = "100%", fig.align = "center"}
cap <- "This figure plots the yearly difference in the share of female-authored readings relative to courses taught by a single male instructor (baseline = 0), for each team configuration. The values are regression coefficients obtained separately for each academic year, with controls for country and field. The outcome is the percentage-point gap in the proportion of female authors cited in the course readings; positive values therefore indicate a larger share of female authors over total authors associated with a given team configuration in that year. The shaded area represents 95% confidence intervals." # nolint

add_figure(file.path(fig_dir, "plot-female-ratio-1.pdf"))
```

Figure \@ref(fig:female-authors) highlights significant and consistent gender association with the percentage of female authors over total authors cited in course readings per year, controlling for academic field and country.[^female-authors] Our results show that all-female courses consistently cite a higher fraction of female authors than all-male ones in the same year, with an effect that decreases over time going from 6% in 2000 to 3% in 2019 (with minor or insignificant differences between courses taught by one vs. two instructors). Mixed-gender courses are somewhat in between, as they cite a larger share of women than all-male courses, but less than all-female courses. Overall, these results indicate a consistent association between gender and cited patterns, suggesting that readings tend to cluster based on the gender of the instructors and mixed-gender teams could promote greater diversity in assigned readings.

[^female-authors]: This analysis excludes syllabi where no readings were matched with the available bibliographic sources and, within the matched readings, excluding the references where the authors' gender remained unidentified.

### Age of Readings (“novelty”)

```{r age-readings, fig.cap = cap, out.width = "100%", fig.align = "center"}
cap <- "This figure illustrates the yearly difference in the percentile rank of the Age of References relative to courses taught by a single male instructor (baseline = 0), for each team configuration. The Age of Reference variable is the syllabus’ year minus the average publication year of the assigned readings. The coefficients are estimated separately for each academic year with controls for country and broad field; positive values indicate that the team configuration assigns newer readings compared with the male-alone baseline in the same year. The shaded area represents 95% confidence intervals." # nolint

add_figure(file.path(fig_dir, "plot-age-of-references-1.pdf"))
```

Figure \@ref(fig:age-readings) illustrates significant gender differences in the publication age of the selected readings.[^age-footnote] Between 2000 and 2019, courses taught by a single female instructor tend to assign newer readings than otherwise similar courses taught by a single man in that year, with a difference that goes from -3% in 2000 to -1% in 2019. Similarly, courses led by two women also assign newer material than those taught by two men, with a decreasing difference. Mixed-gender teams tend to assign newer readings than the baseline in few years, but not consistently in all the years. There are no significant differences between courses taught by two male instructors and the single-male baseline.  Overall, these findings suggest that women tend to assign more novel readings, although this gender difference tend to decrease over time.

[^age-footnote]: This analysis includes only courses with matched bibliographic data. 

## Robustness

To ensure the reliability and robustness of our results, we conducted several additional analyses. In addition to reporting results from separate regressions on subsets of data by year, we also estimated mixed-effects regressions using the full dataset. These results are broadly consistent with those presented in the paper. 

Furthermore, while the reported model specification is our preferred one, we  explored several alternatives. First, we log-transformed our dependent variables -- the age of readings, interdisciplinarity score, and count of female authors --- to explore potential non-linear relationships, particularly relevant for disciplines  with very old readings (e.g., history)

Second, we employed a quasi-poisson mixed-effect model instead of the reported linear mixed-effect regression to better account for non-linearities in the count of female authors. We also trimmed the data to exclude outliers, such as courses with an unusually high number of readings. These changes yield similar results. 

We further tested alternative model specifications such as considering the number of instructors as a numeric variable interacted with the proportion of female instructors to explore a linear association with the dependent variables. While we observe similar results, the linear association could be misleading when extrapolated beyond two instructors. So, we preferred to keep team configurations as fixed effects, as reported here. 

Finally, we further explored specifications distinguishing between male-led and female-led courses, using the gender of the first listed instructor as a proxiy for seniority. Overall, results appear robust to these additional analyses.
