\newpage
\processdelayedfloats

\newpage 
\appendix
\renewcommand{\thefigure}{SI-\arabic{figure}}
\renewcommand{\thesection}{SI}
\renewcommand{\thesubsection}{SI-\arabic{subsection}}

<!-- Set counters -->
\setcounter{figure}{0}
\setcounter{table}{0}
\setcounter{section}{0}
\setcounter{subsection}{0}
\renewcommand{\thetable}{SI-\arabic{table}}

```{r, SI-setup, include = FALSE}
library(yaml)
library(dplyr)
library(tidyr)
library(kableExtra)

create_table <- function(x, ...) {
  kableExtra::kbl(x, ...) %>%
    kableExtra::kable_styling(
      latex_options = c("striped", "hold_position"),
      font_size = 10,
      full_width = FALSE,
      position = "center"
    )
}

```

```{r, data-supporting, cache = TRUE, include = FALSE}
config <- yaml.load_file("config.yml")
data_dir <- config$data_dir

filename_data <- file.path(data_dir, "rds/final.rds")

ds <- readRDS(filename_data) %>%
  filter(nchar(team) < 10) %>%
  select(field, year, team = team_ordered, team_rand, country) %>% 
  pivot_longer(cols = c(team, team_rand),
               names_to = "formation",
               values_to = "composition") %>%
  count(formation, composition, field, year, country)
```

# Supporting Information

## Simulating Gender-Neutral Courses {#si-simulations}

We employed a simlar methodology to that developed by @uzzi2013atypical for the analysis of academic citations. We counted the frequency of gender combinations (male-male, female-male, etc.) of each syllabus per field and academic year. We compared these combinations against those expected by chance, using a gender team composition network. In this network, for a given field, institution, and academic year, we switched all the instructors using a Monte Carlo algorithm. The switching algorithm preserves the total gender counts and the distribution of team size. This ensures that a course with $n$ instructors in the original data will have the same number of instructors in the randomised network. Similarly, an institution with $m$ male instructors and $f$ female instructors teaching in a specific field will have the same number of male and female instructors. The only difference between the randomised and the original data will be the gender composition of the teams.

## Interdiscipline Similarity {#si-interdisciplinarity}

We measured interdisciplinarity using text similarity between syllabi as in @evans2016measuring and @han2023interdisciplinary. We transformed text from course descriptions into "bags of words," with term frequences (TF) normalised using the inverse document frequency (IDF). For each year, we calculated the TF-IDF scores for all syllabi and for 69 academic fields, using concatenated descriptions for the fields. We then computed the weighted average of the cosine similarity, $\cos (i, f)$, between each course \(i\) and field \(f\), where the weight is based on the similarity between field \(f\) and the course's closest field $f_{\max}$. Specifically, $w(f_{\max}, f) = \cos(f_{\max}, f)$, where $f_{\max}$ is the field that has the highest cosine similarity with course \(i\) (i.e., $f_{\max} = \arg\max_f \cos(i, f)$) in that academic year.

Thus, the interdisciplinarity score for course \(i\) is:
$$
  \text{interdisciplinarity score}_i = 1 - \frac{\sum_{f \in (1, 69)} \cos(f_{\max}, f) \cdot \cos(i, f)}{\sum_{f \in (1, 69)} \cos(f_{\max}, f)}. 
$$
To reduce computational costs, we "bootstrap" the field-by-field cosine similarity matrix by using a 10% random sample of syllabi for each academic year. We repeated the sub-sampling procedure ten times and averaged the results.

\clearpage 

## Addtional Tables 

<!-- ### Countries -->

```{r, SI-country}
ds %>%
  filter(formation == "team") %>% 
  count(country, wt = n) %>%
  mutate(
    country = case_when(
      country == "US" ~ "USA",
      country == "GB" ~ "Great Britain",
      country == "CA" ~ "Canada",
      country == "IT" ~ "Italy",
      country == "PL" ~ "Poland",
      country == "NL" ~ "Netherlands",
      country == "DE" ~ "Germany",
      country == "IE" ~ "Ireland",
      country == "PT" ~ "Portugal",
      country == "SE" ~ "Sweden",
      country == "ES" ~ "Spain",
      country == "AT" ~ "Austria",
      country == "DK" ~ "Denmark",
      country == "FR" ~ "France",
      TRUE ~ "Other"
    )
  ) %>%
  mutate(pc = round(100 * n / sum(n), 1)) %>% 
  mutate(n = sprintf("%2.1f", n / 1e3)) %>% 
  create_table(
    caption = "Number of Syllabi per Country",
    col.names = c("Country", "N (thousands)", "%"),
    align = c("l", "r", "r"),
    booktabs = TRUE,
    linesep = "",
    format.args = list(big.mark = " ")
  )
```
<!-- ### Fields -->

```{r SI-fields, echo = FALSE}
ds %>%
  filter(formation == "team") %>% 
  count(field, wt = n) %>%
  mutate(index = row_number(),
         col = ifelse(index %% 2 == 1, "left", "right"),
         group = (index + 1) %/% 2) %>%
  select(-index) %>%
  mutate(pc = round(100 * n / sum(n), 1)) %>% 
  mutate(n = sprintf("%2.1f", n / 1e3)) %>% 
  mutate(across(c(field, n, pc), as.character)) %>% 
  pivot_wider(names_from = col, 
              values_from = c(field, n, pc), 
              values_fill =  "") %>%
  select(field_left, n_left, pc_left, field_right, n_right, pc_right) %>%
  create_table(
    caption = "Number of Syllabi per Field",
    col.names = rep(c("Field", "N (thousands)", "%"), 2),
    align = c("l", "r", "r", "l", "r", "r"),
    booktabs = TRUE,
    linesep = ""
  )
```

<!-- ### Years -->

```{r, SI-years}
ds %>%
  filter(formation == "team") %>%
  mutate(
    year_bc = ifelse(year < 2000, "1999 or older", as.character(year))
  ) %>% 
  count(year_bc, wt = n) %>%
  mutate(pc = round(100 * n / sum(n), 1)) %>%
  mutate(n = sprintf("%2.1f", n / 1e3)) %>% 
  create_table(
    caption = "Number of Syllabi per Year",
    col.names = rep(c("Academic year", "N (thousands)", "%"), 1),
    align = c("l", "r", "r"),
    booktabs = TRUE,
    linesep = ""
  )
```

<!-- ### Gender Composition -->

```{r, SI-composition}
convert_gender_label <- function(original_labels) {
  dplyr::case_match(
    original_labels,
    "f"  ~ "Female only (F)",
    "ff" ~ "Female-Female (FF)",
    "mf" ~ "Mixed Gender (MF/FM)",
    "fm" ~ "Mixed Gender (MF/FM)",
    "mm" ~ "Male-Male (MM)",
    "m"  ~ "Male only (M)",
    .default = "Unknown"
  )
}

ds %>%
  filter(formation == "team") %>%
  count(composition, wt = n) %>%
  arrange(desc(n)) %>%
  mutate(pc = round(100 * n / sum(n), 1),
         n = sprintf("%2.1f", n / 1e3)) %>% 
  filter(nchar(composition) < 3) %>% 
  mutate(composition = convert_gender_label(composition)) %>%
  rename(
    "N (thousands)" = n,
    "Team composition" = composition,
    "%" = pc,
  ) %>%
  create_table(
    caption = "Teaching Team Configurations",
    format.args = list(big.mark = " "),
    align = c("l", "r", "r"),
    booktabs = TRUE,
    linesep = ""
  )
```

