\newpage 
\appendix
\renewcommand{\thefigure}{SI-\arabic{figure}}
\renewcommand{\thesection}{SI}
\renewcommand{\thesubsection}{SI-\arabic{subsection}}
\setcounter{figure}{0}
\setcounter{section}{0}
\setcounter{subsection}{0}

# Supporting Information

## Simulating Gender-Neutral Courses {#si-simulations}

<!-- - We dropped all teams with more than 10 members to minimize the risk of a few ouliers with many members drirving the results. 

- We stablise the estimated proportions of teams with 2 pseudo-observations to make it robust to small samples and avoid that the the proportions to become 0 and 1 for sub-groups with few data as in Agresti et al. (xxx).  -->

We employed a simlar methodology to that developed by @uzzi2013atypical for the analysis of academic citations. We counted the frequency of gender combinations (male-male, female-male, etc.) of each syllabus per field and academic year. We compared these combinations against those expected by chance, using a gender team composition network.

In this network, for a given field, institution, and academic year, we switched all the instructors using a Monte Carlo algorithm. The switching algorithm preserves the total gender counts and the distribution of team size. This ensures that a course with $n$ instructors in the original data will have the same number of instructors in the randomised network. Similarly, an institution with $m$ male instructors and $f$ female instructors teaching in a specific field will have the same number of male and female instructors. The only difference between the randomised and the original data will be the gender composition of the teams.

## Interdiscipline Similarity {#si-interdisciplinarity}

We measured interdisciplinarity using text similarity between syllabi as in @evans2016measuring and @han2023interdisciplinary. We transformed text from course descriptions into "bags of words," with term frequences (TF) normalised using the inverse document frequency (IDF). For each year, we calculated the TF-IDF scores for all syllabi and for 69 academic fields, using concatenated descriptions for the fields. We then computed the weighted average of the cosine similarity, $\cos (i, f)$, between each course \(i\) and field \(f\), where the weight is based on the similarity between field \(f\) and the course's closest field $f_{\max}$. Specifically, $w(f_{\max}, f) = \cos(f_{\max}, f)$, where $f_{\max}$ is the field that has the highest cosine similarity with course \(i\) (i.e., $f_{\max} = \arg\max_f \cos(i, f)$) in that academic year.

Thus, the interdisciplinarity score for course \(i\) is:
$$
  \text{interdisciplinarity score}_i = 1 - \frac{\sum_{f \in (1, 69)} \cos(f_{\max}, f) \cdot \cos(i, f)}{\sum_{f \in (1, 69)} \cos(f_{\max}, f)}. 
$$
To reduce computational costs, we "bootstrap" the field-by-field cosine similarity matrix by using a 10% random sample of syllabi for each academic year. We repeated the sub-sampling procedure ten times and averaged the results.

\clearpage 

## Addtional Figures 

```{r, fig.cap = "Summary of observations for key variables.", fig.align="centering", fig.pos="hp!", , out.width = "80%"}
#include_graphics("figures/summary_obs.pdf")
```

\clearpage

```{r, fig.cap = "Distribution of team types.", fig.align="centering", fig.pos="tp!", out.width="80%"}
#include_graphics("figures/team_distribution.pdf")
```

\clearpage

```{r, fig.cap = "Distribution of syllabi by country over time.", fig.align="centering", fig.pos="tp!", out.width = "80%"}
#include_graphics("figures/countries.pdf")
```

\clearpage

```{r, fig.cap = "Distribution of syllabi by ISCED Fields of Education and Training (ISCED-F 2013) over time.", fig.align="centering", fig.pos="tp!", out.width = "80%"}
#include_graphics("figures/fields.pdf")
```


\clearpage

```{r formation-by-field-aggregated, fig.cap = "This figure highlights the median and interquantile range of the differences in ratios of the simulated versus observed gender composition, disaggregated by the academic fied (ISCED field of education and training).", out.width = "100%", fig.align = "center"}

#knitr::include_graphics("figures/team_formation_field_aggregated.pdf")
```


```{r interdisciplinarity-by-field, fig.cap = "The figure illustrates the differences in interdisciplinarity between female-only (the baseline, set to zero), male-only, and mixed-gender courses. These differences are obtained from separate linear regressions for each academic field and team size, controlling for country and academic year. Interdisciplinarity is measured in percentile ranks per year of the interdisciplinary score, with higher values indicating greater interdisciplinarity scores relative to other courses in the same cohort.", out.width = "100%", fig.align = "center"}

#include_graphics("figures/interdisciplinarity_by_field.pdf")
```


```{r, fig.cap = cap}
cap <- "TBA"
plot(1) 
# Male-led vs Female-led Team Formation.
```