```{r, include = FALSE}
library(dplyr)
library(ggplot2)
library(ggrepel)
library(knitr)

knitr::opts_chunk$set(error = TRUE)

source("../R/theme.R")
theme_set(theme_custom())

term_labels <- c(
    "teamf" = "Woman (F)",
    "teamff" = "Two women (FF)",
    "teamm" = "Man (M)",
    "teammm" = "Two men (MM)",
    "teamfm" = "Mixed-gender (F/M)"
)

plot_coeffs <- function(df) {
    ggplot(
        df,
        aes(
            x = as.numeric(year), 
            y = estimate,
            ymin = conf.low,
            ymax = conf.high
        )
    ) +
        geom_hline(
            aes(yintercept = 0, linetype = "Man alone (M)"),
            color = "red"
        ) +
        scale_linetype_manual(values = "dashed") +
        geom_errorbar(linewidth = 0.25) +
        geom_point() +
        geom_smooth(
            method = "gam", formula = y ~ s(x, bs = "cs"), 
            aes(weight = 1 / std.error^2, color = "Smoothed trend (95% CI)"), 
            linewidth = 0.7
        ) +
        scale_color_brewer(palette = "Set1") + 
        facet_wrap(~term, labeller = labeller(term = term_labels)) +
        labs(
            x = "Year",
        ) + 
        theme(
            axis.title = element_text(face = "bold"),
        )
}
```

```{r, include = FALSE}
results_dir <- file.path("..", "data", "results")
coeffs <- readRDS(file.path(results_dir, "coeffs.rds"))
```

```{r result-setup, include = FALSE}
library(yaml)
config <- yaml.load_file("config.yml")
fig_dir <- file.path(getwd(), config$fig_dir)

knitr::opts_chunk$set(echo = FALSE)

add_figure <- function(path) {
    if (file.exists(path))
        return(knitr::include_graphics(path, auto_pdf = TRUE, rel_path = FALSE))
    warning("File not found: ", path)
    plot(1, main = path)
}
```

# Results

```{r team-trends, fig.cap = cap, out.width = "100%", fig.align = "center"}

cap <- "Evolution of Teaching Teams by Gender Composition and Size. Trends in course gender composition from 2000 to 2019 reveal a steady increase in women's participation (both in solo-taught and two-instructor courses), while the proportion of mixed-gender teams has remained relatively stable over time. (A) Percentage of solo-taught courses by instructor gender over time. (B) Percentage of two-instructor courses by gender composition over time. N = 5.1 million courses. The shaded area represents 95\\% confidence interval of the proportion per academic year." # nolint

readRDS(file.path(results_dir, "fig_evolution.rds"))
```

Figure \@ref(fig:team-trends) illustrates gender composition and team size trends in university courses for single- and two-instructor courses (2000-2019). Panel A shows a decline in male-only single-instructor courses, from approximately 60% in 2000 to 46% in 2019, alongside an increase in female-only single-instructor courses, from roughly 25% to 36%.[^women-academia] Panel B shows male-only two-instructor courses also decreased, from about 8% in the early 2000s to 6% in 2020, while female-only and mixed-gender two-instructor courses remained relatively stable, each accounting for approximately 4% and 3%, respectively. These patterns indicate a consistent shift toward greater gender diversity in course instruction over time, with mixed-gender teaching teams (3%) consistently less common than same-gender teams (10%).

[^women-academia]: This overall trend aligns with the increase in women holding academic positions worldwide. In 2000, women comprised approximately 35% of the global academic staff. By 2022, this figure had risen to about 44%, according to data from the World Bank. The data are available at: https://data.worldbank.org/indicator/SE.TER.TCHR.FE.ZS

```{r montecarlo, fig.cap = cap, out.width = "100%", fig.align = "center", echo = FALSE}
cap <- "Comparison of gender composition between actual courses and courses simulated with Monte Carlo. Simulations ensure gender neutrality while keeping constant institutional constraints (i.e., number of teams by size, institution, academic year, and 69 academic fields). Simulated mixed-gender teams consistently exceed observed proportions, highlighting persistent barriers to forming gender-diverse teaching teams." # nolint

readRDS(file.path(results_dir, "fig_montecarlo.rds"))
```

To assess whether gender imbalances in co-teaching teams persist after accounting for institutional constraints across fields and institutions, Figure \@ref(fig:montecarlo) compares observed teaching team compositions (2000–2019) with those generated by a gender-neutral Monte Carlo simulation. This model randomly shuffles instructors while preserving the overall distribution of course loads per year by institution and academic field but ignoring gender-based preferences or constraints (see Section \@ref(sec:methods)).

The results reveal that actual data systematically deviate from simulated expectations: same-gender teams, such as female-female (FF) and male-male (MM) courses, occur *more* frequently in actual data than in simulations by 66% and 20%, respectively, in 2019. By contrast, mixed-gender collaborations (MF/FM) are consistently *underrepresented* in the actual data (3%) compared to simulations (6%). These findings suggest that significant social or institutional dynamics---beyond chance---reinforce same-gender pairings, especially among women. Such dynamics may include gender-based homophily, mentoring networks, or within-department barriers or assignment practices.

```{r montecarlo-by-field, fig.cap = cap, out.width = "100%", fig.align = "center"}
cap <- "This figure disaggregates the comparison of proportions between actual (green) and simulated (orange) mixed-gender (MF/FM) courses by academic field. The shaded bar indicates a 95% confidence level of the proportion." # nolint

readRDS(file.path(results_dir, "fig_montecarlo_by_field.rds"))
```

Figure \@ref(fig:montecarlo-by-field) disaggregates the comparison between actual and simulated mixed-gender (MF/FM) courses by academic field. Nearly all fields exhibit a stark underrepresentation of mixed-gender teams (relative to expectations) by a factor of two to four times. For instance, in Medicine (Health and Welfare), the actual share of mixed-gender collaborations is 4.8% compared to an expected 9.1%, indicating a significant imbalance. Similar gaps are evident in Law (3.8% actual vs. 6.4% simulated) and Linguistics (4.3% vs. 6.7%), underscoring the pervasiveness of barriers to mixed-gender teams across various academic fields. Remarkably, we observe no differences in these patterns between male-dominated and female-dominated fields (e.g., Engineering or Accounting vs. Nursing or Chemistry), suggesting that the underrepresentation of mixed-gender teams may be a structural feature rather than one driven by field-specific gender imbalances.

```{r montecarlo-by-cntry, fig.cap = cap}
cap <- "This figure disaggregates the comparison of proportions between actual (green) and simulated (orange) mixed-gender courses (MF/FM) by country.  Simulated mixed-gender teams consistently exceed observed proportions in all countries. EU countries are Austria, Denmark, France, Germany, Ireland, Italy, Netherlands, Poland, Portugal, Spain and Sweden." # nolint

readRDS(file.path(results_dir, "fig_montecarlo_by_country.rds"))
```

Figure \@ref(fig:montecarlo-by-cntry) illustrates that the proportion of mixed-gender teams in the simulations consistently exceeds the observed proportions across various geographic regions, despite notable variations in the magnitude of these differences. In 2019, for example, mixed-gender teams are relatively rare in Great Britain (1%) compared to Canada (4%). However, in the simulations, the proportion of mixed-gender teams in both countries increases substantially, 1% to 3% and 4% to 5%, respectively. In 2019, the most considerable discrepancy between simulated and observed proportions is in EU countries, where it ranges from 5% to 12%. Overall, these patterns underscore the robustness of the findings across diverse geographic contexts.

## Analysis Pre- and Post- COVID-19 in Europe

Using a recent dataset of 2.2 million European syllabi from 2017–2022, we examine how the composition of teaching teams changed before and after the COVID-19 pandemic. Across countries and academic fields, the total number of courses declined sharply in the post-COVID period (Figure \@ref(fig:covid-trend), in SI). However, regression analyses reveal that the magnitude of this decline differed substantially depending on the gender composition of instructors (Figure \@ref(fig:covid-coeffs)). After COVID-19, the share of single-instructor courses fell by more than 26%, a significant and substantial drop relative to the pre-pandemic period. Within multi-instructor teams, the proportion of mixed-gender collaborations decreased by up to 10%. Finally, among single-instructor courses, the share of women instructors declined by approximately 1% relative to men. Taken together, these results highlight that single-instructor courses declined more sharply than collaborative ones, and that among collaborative teams, mixed-gender groups were disproportionately affected by the pandemic. Women were also slightly more likely to reduce their participation, although this effect is modest in size.

```{r load-covid, include = FALSE}
fig_covid <- readRDS(file.path(results_dir, "fig_covid.rds"))
```

```{r covid-coeffs, fig.cap = cap, fig.width = 5, fig.asp = .33}
cap <- paste(
    "**Team composition post-Covid-19 change**. Coefficients and 95% confidence intervals for post-covid period obtained from the following Binomial regressions: (a) probability of single teams being taught by a woman vs man; (b) probability of teams being composed by one instructor vs multiple (teaching collaboration); (c) probability of teaching teams being mixed-gender vs. same-gender collaborations. All regressions are based on 2.2M syllabi from institutions in thirty-three European countries (excluding teams with three or more collaborations, and authors of unknown gender)." # nolint
)
fig_covid$coeffs + 
    labs(y = NULL, x = "Post-covid Difference")
```

Figure \@ref(fig:covid-simulations) shows the results of our Montecarlo simulations for this dataset. As highlighted in the previous analysis, there are disproportionately fewer mixed-gender teams than one would expect under gender-neutral matching. However, post-Covid, the gap appears to widen substantially in 2022. This pattern indicates that the pandemic, particularly the lockdown measures implemented across Europe, may have significantly hindered the formation of mixed-gender teaching teams.

```{r covid-simulations, fig.cap = cap, fig.width = 7, fig.asp = .45}
cap <- paste(
    "**Post-Pandemic Team Formation**. Simulated gender-neutral vs observed teaching teams for 2.2M syllabi from institutions in thirty-three European countries. Covid-19 period highlighted (2020-2021). This shows that the gap between mixed-gender and same-gender collaborations widens after Covid." # nolint
)
fig_covid$simulations + labs(title = NULL, subtitle = NULL)
```

## Associations Between Teams and Course Content

To examine how teaching team size and gender composition are associated with important aspects of course materials, we analyse three key metrics: the percentile rank of the interdisciplinarity score per year ("interdisciplinarity"), the average publication year of the readings ("age of readings"), and the share of female authors cited in the assigned readings ("share of female authors"). As previously discussed, we employ linear mixed-effects regressions to examine these relationships while accounting for unobserved differences across academic fields, institutions, time, and other relevant covariates.

### Interdisciplinarity

```{r interdisciplinarity, fig.cap = cap}
cap <- paste(
    "**Interdisciplinarity**. Estimates and 95% confidence intervals for differences in Interdisciplinarity between courses taught by men alone (reference) and courses taught by other instructor teams. Interdisciplinarity is the percentile rank, within a given year, of a syllabus' interdisciplinarity score based on its textual description relative to other syllabi. Estimates are from separate mixed-effects regressions for each academic year, controlling for country, field, institution, course level, probability that the document is a syllabus, and number of readings, with weights proportional to the square root of the number of readings. Smoothed curves show local estimates averaged across years, weighted by the precision (standard errors) of the estimates." # nolint
)
coeffs %>% 
    filter(model == "fit_interdisc", grepl("team", term)) %>%
    plot_coeffs() +
    labs(y = "Interdisciplinarity Difference")
```

Figure \@ref(fig:interdisciplinarity) illustrates the evolving variation in interdisciplinarity across different team configurations, accounting for unobserved effects for academic field and country. When teaching alone, our analysis reveals little differences in interdisciplinarity between female and male instructors. However, in recent years, female instructors have increasingly taught significantly more interdisciplinary solo courses. By contrast, mixed-gender consistently exhibit higher interdisciplinarity compared to all-male teams. As the figure illustrates, in most years, the difference between mixed-gender and male-alone courses is statistically significant (p < 0.05), and teams with two male instructors tend to be less interdisciplinary than solo-male courses. Notably, we find no significant evidence of a difference between mixed-gender and teams with two female instructors. These results point to a robust association between gender diversity and interdisciplinarity.

### Citing Women Authors

```{r female-authors, fig.cap = cap}
cap <- paste(
     "**Female-Authored Radings Ratio**. Estimates and 95% confidence intervals for differences in Female-authored radings ratio between courses taught by men alone (reference) and courses taught by other instructor teams. This variable is the proportion of female authors cited in the course readings. Estimates are from separate mixed-effects regressions for each academic year, controlling for country, field, institution, course level, probability that the document is a syllabus, and number of readings, with weights proportional to the square root of the number of readings. Smoothed curves show local estimates averaged across years, weighted by the precision (standard errors) of the estimates." # nolint
)
coeffs %>%
    filter(model == "fit_women", grepl("team", term)) %>%
    plot_coeffs() +
    scale_y_continuous(labels = scales::percent) +
    labs(
        y = "Female-Authored Readings Ratio Difference"
    )
```

Figure \@ref(fig:female-authors) highlights significant and persistent gender association with the percentage of female authors over total authors cited in course readings per year, controlling for academic field and country.[^female-authors] All-female courses consistently cite a higher fraction of female authors than all-male ones in the same year, with an effect that decreases over time, going from 8% in 2000 to 4% in 2019, with minor or insignificant differences between courses taught by one vs. two instructors. Mixed-gender courses are in between, as they cite a larger share of women than all-male courses but less than all-female courses. Overall, these results indicate a consistent association between gender and cited patterns, suggesting that readings tend to cluster based on the gender of the instructors, and mixed-gender teams could promote greater diversity in assigned readings.

[^female-authors]: This analysis excludes syllabi with no matched readings with the available bibliographic sources and, within the matched readings, excludes the references where the authors' gender remained unidentified.

### Teaching Conventionality, Age of Readings, and Scientific Novelty

Figure \@ref(fig:age-readings) reveals persistent temporal trends and significant gender differences in the publication age of the selected readings, even after accounting for observable differences across multiple dimensions, including course level, academic field, institution, and country. Course taught by a single female instructor tend to assign more recent readings than comparable courses taught by a single man instructor (the reference category), with the difference going from -3% in 2000 to -1% in 2019. Similarly, courses led by two women also assign newer material than those taught by two men. Mixed-gender teams generally assign more recent readings than single-male instructors, although less than all-female teams of equal size. No significant differences emerge between courses taught by two male instructors and the single-male baseline.  Overall, these findings indicate that the presence of women in teaching teams, either alone or with a male co-instructor, is associated with more recent readings, although the gender effect appears to decrease over time.

```{r age-readings, fig.cap = cap, fig.width = 7, fig.asp = 0.75}
cap <- paste(
    "**Age of Readings by Team**. Estimates and 95% confidence intervals for differences in Age of Readings between courses taught by men alone (reference) and courses taught by other instructor teams. Age of Readings is the percentile rank, within a given year, of a syllabus' average publication age of the readings. Estimates are from separate mixed-effects regressions for each academic year, controlling for country, field, institution, course level, probability that the document is a syllabus, and number of readings, with weights proportional to the square root of the number of readings. Smoothed curves show local estimates averaged across years, weighted by the precision (standard errors) of the estimates." # nolint
)
coeffs %>%
    filter(model == "fit_age_readings", grepl("team", term)) %>%
    plot_coeffs() +
    labs(y = "Age of Readings Difference\n(Percentile Rank Per Year)")
```

Figure \@ref(fig:conventionality) illustrates significant gender differences in the readings' conventionality, which is the median citations of syllabus readings from references in other syllabi. Team-taught courses tend to be more conventional than single-instructor courses, as both mixed-gender and all-male teams exhibit significantly higher conventionality than the baseline, while the difference with two-women teams is insignificant. Courses taught by women alone are significantly less conventional than those taught by men alone, but the difference is small (less than 1 pp). Taken together, this evidence indicates a strong association between reading conventionality and teaching in teams with more than one instructor, which is partly offset by the presence of women.

```{r conventionality, fig.cap = cap, fig.width = 7, fig.asp = 0.75}
cap <- paste(
    "**Readings Conventionality by Team**. Estimates and 95% confidence intervals for differences in Teaching Conventionality between courses taught by men alone (reference) and courses taught by other instructor teams. Teaching Conventionality is the percentile rank, within a given year, of a syllabus' median citations of readings, based on references from other syllabi. Estimates are from separate mixed-effects regressions for each academic year, controlling for country, field, institution, course level, probability that the document is a syllabus, and number of readings, with weights proportional to the square root of the number of readings. Smoothed curves show local estimates averaged across years, weighted by the precision (standard errors) of the estimates." # nolint
)
coeffs %>%
    filter(model == "fit_conventional", grepl("team", term)) %>% 
    plot_coeffs() +
    labs(y = "Readings Conventionality Difference")
```

Figure \@ref(fig:atypicality) presents the results of our analysis of the scientific novelty of the readings, based on approximately 400,000 syllabi for which we obtained Uzzi et al. [@uzzi2013atypical]'s atypicality measure precomputed by SciSciNet (v2) [@lin2023sciscinet].  Controlling for field, course level, and other relevant characteristics, the results indicate that readings produced by mixed-gender teams are significantly less at the frontier of knowledge.^[
    This regression also inclused the age of readings as a control for the novelty at time of publication of the article.
] Similarly, women-led teams show a negative association compared to man-teaching-alone courses, with the effect being stronger in courses taught by two women. In other words, men-only teams exhibit slightly higher atypicality in our dataset than any other team configuration. 

```{r atypicality, fig.cap = cap, fig.width = 7, fig.asp = 0.75}
cap <- paste(
    "**Scientific Novelty by Team**. Estimates and 95% confidence intervals for differences in Scientific Novelty between courses taught by men alone (reference) and courses taught by other instructor teams. Scientific Novelty is the percentile rank, within a given year, of a syllabus' median Uzzi et al's atypicality of readings, based on the combination of references in the readings. Estimates are from separate mixed-effects regressions for each academic year, controlling for country, field, institution, course level, probability that the document is a syllabus, and number of readings, with weights proportional to the square root of the number of readings. Smoothed curves show local estimates averaged across years, weighted by the precision (standard errors) of the estimates." # nolint
)
coeffs %>%
    filter(model == "fit_atypical", grepl("team", term)) %>% 
    plot_coeffs() +
    labs( y= "Readings Atypicality Difference")
```

## Heterogeneity

Figure \@ref(fig:stem) highlights several key differences between STEM and non-STEM fields, based on separate models estimated for STEM fields compared against results from the full sample.^[
    We converted CIP fields into ISCED-F broad categories and considered STEM the following: (1) Natural Sciences, Mathematics, and Statistics,
    (2) Information and Communication Technologies (ICTs), and (3) Engineering, Manufacturing, and Construction.
]. The gap in female-author citations associated with women instructors remains consistent but is significantly smaller in STEM fields, roughly half the size observed in all fields. The positive association between mixed-gender teams and conventional readings is larger in STEM fields. Conversely, the gap in atypicality widens for women teaching alone, although the effect has been diminishing over time. In contrast, the gap in the age of assigned readings between women and men teaching alone is smaller in STEM. With respect to interdisciplinarity, the estimated differences are small and not consistently significant across years. As a result, the evidence for heterogeneity in interdisciplinarity within STEM is inconclusive.

```{r stem, fig.cap = cap, fig.asp = 1, fig.width = 9}
cap <- paste(
    "**Heterogeneity within STEM Fields**. Estimates and 95% confidence intervals for differences in (1) interdisciplinarity (2) readings' conventionality (3) age of readings (4) female-authored ratio and (5) readings' atypicality. Estimates are from separate mixed-effects regressions for STEM fields vs the full sample, in each academic year, controlling for country, field, institution, course level, probability that the document is a syllabus, and number of readings, with weights proportional to the square root of the number of readings. Smoothed curves show local estimates averaged across years, weighted by the precision (standard errors) of the estimates." #nolint
)

library(patchwork)
names <- c("interd", "conv", "age", "women", "atyp")
plots <- lapply(names, \(x) {
    coeffs %>%
        filter(
            grepl("team", term), 
            grepl(x, model),
            !grepl("_us", model)
        ) %>%
        mutate(
            depvar = gsub("fit_([^_]+)(_.*)?", "\\1", model),
            stem = ifelse(grepl("stem", model), "STEM", "All"),
        ) %>%
        ggplot(
            aes(
                x = as.numeric(year),
                y = estimate,
                color = stem,
            )
        ) +
        geom_hline(
            aes(yintercept = 0, linetype = "Man alone (M)"),
            color = "red"
        ) +
        scale_linetype_manual(values = "dashed") +
        facet_grid(depvar ~ term, labeller = labeller(term = term_labels)) +
        geom_smooth(
            method = "gam",
            formula = y ~ s(x, bs = "cs"),
            aes(weight = 1 / std.error^2),
        ) + 
        theme(axis.title = element_blank())
})
plots[[1]] / plots[[2]] / plots[[3]] / plots[[4]] / plots[[5]] + 
    plot_annotation(tag_levels = "A") + 
    plot_layout(guides = "collect")
```

We have also explored heterogeneity across countries by fitting separate models for the US and comparing the results against the estimates from the full sample. The analysis revealed no signficant differences for the US-based regressions (Figure \@ref(fig:hetero-us-plots), SI). Therefore, our study provides no evidence of cross-country differences in the observed associations.

```{r hetero-us, fig.asp = 1, fig.width = 9, include = FALSE}
library(patchwork)
names <- c("interd", "conv", "age", "women", "atyp")
plots <- lapply(names, \(x) {
    coeffs %>%
        filter(
            grepl("team", term), 
            grepl(x, model), 
            !grepl("_stem", model)
        ) %>%
        mutate(
            depvar = gsub("fit_([^_]+)(_.*)?", "\\1", model),
            us = ifelse(grepl("us", model), "US", "All"),
        ) %>%
        ggplot(
            aes(
                x = as.numeric(year),
                y = estimate,
                color = us,
            )
        ) +
        geom_hline(
            aes(yintercept = 0, linetype = "Man alone (M)"),
            color = "red"
        ) +
        scale_linetype_manual(values = "dashed") +
        facet_grid(depvar ~ term, labeller = labeller(term = term_labels)) +
        geom_smooth(
            method = "gam",
            formula = y ~ s(x, bs = "cs"),
            aes(weight = 1 / std.error^2),
        ) + 
        theme(axis.title = element_blank())
})
plots_us <- plots[[1]] / plots[[2]] / plots[[3]] / plots[[4]] / plots[[5]] + 
    plot_annotation(tag_levels = "A") +
    plot_layout(guides = "collect")
```

## Robustness

The robustness checks conducted, as detailed in the Supplementary Information (Section \@ref(robustness)), show that the main results are consistent across multiple alternative analyses. These included mixed-effects regressions on the full dataset, log-transformations of key dependent variables, quasi-Poisson models for counts of female authors, trimming outliers, and testing different specifications related to instructor composition and seniority. Across all these variations, the findings remained broadly stable, demonstrating that the results are not sensitive to specific modeling choices or data subsets.