---
title: |
  Gender in Teaching:
  Insights from Five Million Syllabi on Collaboration, Interdisciplinarity, and Reading Selections

author:
- correspondence: yes
  email: 'andrea.blasco@ec.europa.eu'
  id: Andrea Blasco
  institute:
  - jrc
  name: Andrea Blasco

- id: Elisa Orlandi
  institute:
  - jrc
  - dulbea
  name: Elisa Orlandi

institute:
- id: jrc
  name: European Commission, Joint Research Centre
- id: dulbea
  name: Université libre de Bruxelles


abstract: |
  This study examines how academic teaching teams form and how their gender composition is associated with course design and gender disparities in academic recognition. Analysing over five million syllabi from higher education in fifteen countries, we document a persistent underrepresentation of mixed-gender co-teaching teams, which occur 50% less frequently than expected under gender-neutral team formation. We examine how team gender composition relates to interdisciplinarity, the novelty of reading materials, and the reppresentation of women authors in assigned readings. Courses taught by mixed-gender teams are significantly more interdisciplinary than those taught by same-gender teams. Male-only teams cite substantially fewer female authors than female-only teams, and this gap persists in mixed-gender teams. Courses taught by solo female instructors or teams including at least one woman tend to incorporate more recent reading materials. However, mixed-gender teams tend to rely more on conventional readings and fewer readings at the research frontier. These findings reveal persistent gendered patterns in teaching collaboration and course content, while highlighting the relationship between team composition, interdiscipliniarity, and the representation of women authors in academic teaching.
date: >
  This version: `r format(Sys.Date(), "%B %d, %Y")`
always_allow_html: true
fontsize: 12pt
linestretch: 1.5
colorlinks: true
link-citations: true
bibliography: [refs.bib]
csl: [assets/pnas.csl]
geometry: margin=1in
lang: en-UK
mainfont: Times New Roman
---

<small>
JEL Classification:  I21, I23, J16
</small>

<small>
Keywords: education | gender bias | teams | interdisciplinarity | novelty | syllabi
</small>

\newpage

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    echo = FALSE,
    message = FALSE,
    warning = FALSE,
    error = TRUE
)
sections <- list.files("sections", pattern = "*.Rmd", full.names = TRUE)
```

```{r sections, child = sections}
```

\clearpage
\ifcsname processdelayedfloats\endcsname
\processdelayedfloats
\fi
\clearpage

# Acknowledgements {-}

The contents of this publication do not necessarily reflect the position or opinion of the European Commission or the Joint Research Centre.

# Data Availability Statement {-}

All the datasets generated during the current study are available online in the [name of repository], [persistent web link or DOI]. The raw data will be available from the corresponding author upon reasonable request and if the request meets the contractual agreements with OpenSyllabus.

All the scripts and code used in the analyses is available at [GitHub link or DOI]. This includes two repositories one with all the Python scripts used to compute the interdisciplinarity score [LINK or DOI] and another with Rscripts to analyse the data and create the figures [LINK or DOI].

# References {-}

::: {#refs}
:::

```{r appendix, child = "50-supporting.Rmd", eval = FALSE}
```