# Results

## Team formation

```{r trends, fig.cap = cap, fig.align = "center", echo = FALSE}
cap <- "Trends in the gender composition of university courses, 2000-2019. Left, percentage of solo-taught courses by instructor gender. Right, percentage of two-instructor courses by gender composition. The proportion of mixed-gender teams remains relatively stable. N = 5.1 million courses." # nolint

knitr::include_graphics(
    here::here(
        "output/01_plot_montecarlo_analysis/01_annual_trends.pdf"
    )
)
```

Figure \@ref(fig:trends) illustrates changes in the gender composition of university courses between 2000 and 2019. In single-instructor courses (left panel), the share of female-only courses increased from approximately 30% to 45% between 2000 and 2019, aligned with the increase in women holding academic positions worldwide.[^women-academia]  Likewise the share of female-only two-instructor courses increased from 20% in 2000 to 35% in 2019. At the same time, the share of mixed-gender courses remained relatively stable at 20% over the same period.

[^women-academia]: In 2000, women comprised approximately 35% of the global academic staff. By 2022, this figure had risen to about 44%, according to data from the World Bank. The data are available at: https://data.worldbank.org/indicator/SE.TER.TCHR.FE.ZS

```{r montecarlo, fig.cap = cap}
cap <- "Comparison of gender composition between actual courses and courses simulated with Monte Carlo. Simulations ensure gender neutrality while keeping constant institutional constraints (i.e., number of teams by size, institution, academic year, and 69 academic fields). Simulated mixed-gender teams consistently exceed observed proportions, highlighting persistent barriers to forming gender-diverse teaching teams." # nolint

knitr::include_graphics(
    here::here(
        "output/01_plot_montecarlo_analysis/02_obs_vs_simul.pdf"
    )
)
```

Figure \@ref(fig:montecarlo) compares the observed trends with those generated by a gender-neutral Monte Carlo simulation (Sec. \@ref(sec:methods)), showing that mixed-gender courses are consistently *underrepresented* in the actual data compared to simulations, while same-gender courses are consistently *over-represented*. Specifically, the observed share of mixed-gender teams led by a male instructor (male-female) is 8.5 percentage points lower than in the simulations. Similarly, the share of female-male teams is 10.6 percentage points lower than in the simulations. These differences remain remarkably stable after 2005, despite the growing incidence of women-led courses in academia. Taken together, these findings provide evidence that is consistent with either persistent barriers or gender-based preferences that may discourage mixed-gender collaboration.

```{r montecarlo-by-field, fig.cap = cap}
cap <- "This figure disaggregates the comparison of proportions between actual (green) and simulated (orange) mixed-gender (MF/FM) courses by academic field. The shaded bar indicates a 95% confidence level of the proportion." # nolint

knitr::include_graphics(
    here::here(
        "output/01_plot_montecarlo_analysis/03_by_field.pdf"
    )
)
```


Figure \@ref(fig:montecarlo-by-field) disaggregates the above comparison by academic field, showing the underrepresentation of mixed teams is consistent across fields, although its size varies substantially. The largest shortfalls is in Business, Administration and Law. This pattern is driven primarily by an under-representation of female–male (`fm`) teams, which are 6.7–14.6 percentage points less frequent than simulated across fields. Conversely, same-sex teams are more prevalent than simulated: female–female (`ff`) teams exceed simulated levels by approximately 6.9–14.5 percentage points, while male–male (`mm`) teams are 5.9–13.4 percentage points higher. Overall, the results indicate a systematic tendency towards segregation by sex in two-instructor teams across fields. 
<!-- No evidence of sorting by field by women -->

```{r montecarlo-by-cntry, fig.cap = cap, echo = FALSE}
cap <- "This figure disaggregates the comparison of proportions between actual (green) and simulated (orange) mixed-gender courses (MF/FM) by country.  Simulated mixed-gender teams consistently exceed observed proportions in all countries. EU countries are Austria, Denmark, France, Germany, Ireland, Italy, Netherlands, Poland, Portugal, Spain and Sweden." # nolint

knitr::include_graphics(
    here::here(
        "output/01_plot_montecarlo_analysis/04_by_country.pdf"
    )
)
```

Figure \@ref(fig:montecarlo-by-cntry) illustrates that the observed proportion of mixed-gender teams consistently fall short of what is expected in the simulations across the US, EU member states, Canada, Great Britain, and a few other countries. Despite variations in the magnitude of these differences, these patterns underscore the robustness of the findings across diverse geographic contexts.

<!-- In 2019, for example, mixed-gender teams are relatively rare in Great Britain (1%) compared to Canada (4%). However, in the simulations, the proportion of mixed-gender teams in both countries increases substantially, 1% to 3% and 4% to 5%, respectively. In 2019, the most considerable discrepancy between simulated and observed proportions is in EU countries, where it ranges from 5% to 12%. Overall, these patterns underscore the robustness of the findings across diverse geographic contexts. -->

### Pre- and Post- COVID-19 in Europe

Using a recent dataset of 2.2 million European syllabi from 2017–2022, we examine how the composition of teaching teams changed before and after the COVID-19 pandemic. Across countries and academic fields, the total number of courses declined sharply in the post-COVID period (Figure \@ref(fig:covid-trend), in SI). However, regression analyses reveal that the magnitude of this decline differed substantially depending on the gender composition of instructors (Figure \@ref(fig:covid-coeffs)). After COVID-19, the share of single-instructor courses fell by more than 26%, a significant and substantial drop relative to the pre-pandemic period. Within multi-instructor teams, the proportion of mixed-gender collaborations decreased by up to 10%. Finally, among single-instructor courses, the share of women instructors declined by approximately 1% relative to men. Taken together, these results highlight that single-instructor courses declined more sharply than collaborative ones, and that among collaborative teams, mixed-gender groups were disproportionately affected by the pandemic. Women were also slightly more likely to reduce their participation, although this effect is modest in size.

```{r covid, fig.cap = cap}
cap <- paste(
    "**Team composition post-Covid-19 change**. Coefficients and 95% confidence intervals for post-covid period obtained from the following Binomial regressions: (a) probability of single teams being taught by a woman vs man; (b) probability of teams being composed by one instructor vs multiple (teaching collaboration); (c) probability of teaching teams being mixed-gender vs. same-gender collaborations. All regressions are based on 2.2M syllabi from institutions in thirty-three European countries (excluding teams with three or more collaborations, and authors of unknown gender)." # nolint
)

knitr::include_graphics(
    here::here(
        "output", "02_plot_covid_analysis", 
        "01_covid_coeffs.pdf"
    )
)
```

Figure \@ref(fig:covid-simulations) shows the results of our Montecarlo simulations for this dataset. As highlighted in the previous analysis, there are disproportionately fewer mixed-gender teams than one would expect under gender-neutral matching. However, post-Covid, the gap appears to widen substantially in 2022. This pattern indicates that the pandemic, particularly the lockdown measures implemented across Europe, may have significantly hindered the formation of mixed-gender teaching teams.

```{r covid-simulations, fig.cap = cap, fig.width = 7, fig.asp = .45}
cap <- paste(
    "**Post-Pandemic Team Formation**. Simulated gender-neutral vs observed teaching teams for 2.2M syllabi from institutions in thirty-three European countries. Covid-19 period highlighted (2020-2021). This shows that the gap between mixed-gender and same-gender collaborations widens after Covid." # nolint
)
knitr::include_graphics(
    here::here(
        "output/02_plot_covid_analysis/01_covid_sims.pdf"
    )
)
```

## Associations with Course Materials

Figure \@ref(fig:associations) summarises the associations between teaching team composition with five characteristics of course materials: interdisciplinarity, readings age, the share of female authors among cited works, conventionality and atypicality of the readings. Each panel reports estimates along with 95% confidence intervals (CI) from separate linear mixed-effects regressions across all academic years, accounting for unobserved heterogeneity across academic fields and institutions, temporal variation, and other observable course characteristics.

```{r associations, fig.cap = cap}
cap <- "TBA"
knitr::include_graphics(
    here::here(
        "output/04_make_plots_v2/main_forest_plot.pdf"
    )
)
```

Overall, the results point to a consistent assocation between the gender composition of teaching teams and several dimensions of course materials. First, relative to single-male courses, the percentile rank of the **year-specific interdisciplinarity score** was highest for mixed-gender courses (1.2, [1.0, 1.5]), followed by female–female courses (1.1, [0.9, 1.3]) and single-female courses (0.2, [0.1, 0.2]). In contrast, the difference for male–male courses was small and statistically uncertain (0.1, [−0.1, 0.2]). This suggests that the presence of women in teaching teams is significantly associated with the course's interdisciplinarity ranking within the academic year. 

Second, the **fraction of women authors cited in readings** was 7.6 percentage points higher in female–female courses [7.3, 7.8] and 7.0 percentage points higher in single-female courses [6.9, 7.1], relative to single-male courses. It was also 3.0 percentage point higher in mixed-gender courses [2.7, 3.2], but slightly lower in male–male courses (−0.3, [−0.5, −0.2]). This evidence suggests the presence of women in teaching teams is strongly associated with gender represenation of women authors in course readings.

Third, in terms of the readings' novelty or recency, **Uzzi’s conventionality**, expressed as a within-year percentile rank, was lower in single-female courses (−0.1, [−0.2, −0.1]) and female–female courses (−0.4, [−0.6, −0.3]) but higher in male–male courses (0.9, [0.7, 1.0]) and mixed-gender courses (0.9, [0.7, 1.1]), relative to single-male courses. **Atypicality**, likewise measured as a within-year percentile rank, was lower across all course types, with the largest difference observed for mixed-gender courses (−1.6, [−2.1, −1.0]), followed by female–female (−0.6, [−1.1, −0.1]), male–male (−0.5, [−0.9, −0.1]), and single-female courses (−0.4, [−0.6, −0.2]). Finally, the **age of readings**, also expressed as a within-year percentile rank, was lower in single-female (−1.3, [−1.4, −1.3]), female–female (−1.6, [−1.8, −1.4]), and mixed courses (−0.5, [−0.7, −0.3]), while there was no clear difference for male–male courses (−0.0, [−0.2, 0.1]). Overall, this evidence suggests that the presence of male instructors is associated with readings that are both more conventional and more atypical. In contrast, the presence of women instructors is associated with more recent readings, as indicated by their lower age. Taken together, these findings point to a complex relationship between instructor gender and the novelty and recency of the assigned readings.

```{r temporal-stability, fig.cap = cap}
cap <- paste(
  "**Temporal stability of associations between teaching-team composition and course materials.**",
  "Estimates and 95% confidence intervals for differences relative to courses taught by a single male instructor.",
  "Points show estimates from separate mixed-effects regressions for each academic year; smoothed curves show local estimates averaged across years and weighted by the precision of the yearly estimates."
)

knitr::include_graphics(
    here::here(
        "output/04_make_plots_v2/supp_temporal_stability.pdf"
    )
)
```

To examine whether these associations are stable over time, Figure \@ref(fig:temporal-stability) reports the corresponding year-specific estimates and their smoothed trends. The figure shows that the main differences are generally persistent across the period, while also revealing some temporal variation in their magnitude. In particular, the associations with interdisciplinarity and the representation of women among cited authors show relatively clear changes over time, whereas the differences in conventionality and atypicality are more stable. The temporal patterns are discussed in greater detail alongside the corresponding outcomes below.

## Heterogeneity

Figure \@ref(fig:stem) highlights several key differences between STEM and non-STEM fields, based on separate models estimated for STEM fields compared against results from the full sample.^[
    We converted CIP fields into ISCED-F broad categories and considered STEM the following: (1) Natural Sciences, Mathematics, and Statistics,
    (2) Information and Communication Technologies (ICTs), and (3) Engineering, Manufacturing, and Construction.
]. The gap in female-author citations associated with women instructors remains consistent but is significantly smaller in STEM fields, roughly half the size observed in all fields. The positive association between mixed-gender teams and conventional readings is larger in STEM fields. Conversely, the gap in atypicality widens for women teaching alone, although the effect has been diminishing over time. In contrast, the gap in the age of assigned readings between women and men teaching alone is smaller in STEM. With respect to interdisciplinarity, the estimated differences are small and not consistently significant across years. As a result, the evidence for heterogeneity in interdisciplinarity within STEM is inconclusive.

```{r stem, fig.cap = cap, fig.asp = 1, fig.width = 9, eval = FALSE}
cap <- paste(
    "**Heterogeneity within STEM Fields**. Estimates and 95% confidence intervals for differences in (1) interdisciplinarity (2) readings' conventionality (3) age of readings (4) female-authored ratio and (5) readings' atypicality. Estimates are from separate mixed-effects regressions for STEM fields vs the full sample, in each academic year, controlling for country, field, institution, course level, probability that the document is a syllabus, and number of readings, with weights proportional to the square root of the number of readings. Smoothed curves show local estimates averaged across years, weighted by the precision (standard errors) of the estimates." #nolint
)

library(patchwork)
names <- c("interd", "conv", "age", "women", "atyp")
plots <- lapply(names, \(x) {
    coeffs %>%
        filter(
            grepl("team", term), 
            grepl(x, model),
            !grepl("_us", model)
        ) %>%
        mutate(
            depvar = gsub("fit_([^_]+)(_.*)?", "\\1", model),
            stem = ifelse(grepl("stem", model), "STEM", "All"),
        ) %>%
        ggplot(
            aes(
                x = as.numeric(year),
                y = estimate,
                color = stem,
            )
        ) +
        geom_hline(
            aes(yintercept = 0, linetype = "Man alone (M)"),
            color = "red"
        ) +
        scale_linetype_manual(values = "dashed") +
        facet_grid(depvar ~ term, labeller = labeller(term = term_labels)) +
        geom_smooth(
            method = "gam",
            formula = y ~ s(x, bs = "cs"),
            aes(weight = 1 / std.error^2),
        ) + 
        theme(axis.title = element_blank())
})
plots[[1]] / plots[[2]] / plots[[3]] / plots[[4]] / plots[[5]] + 
    plot_annotation(tag_levels = "A") + 
    plot_layout(guides = "collect")
```

We have also explored heterogeneity across countries by fitting separate models for the US and comparing the results against the estimates from the full sample. The analysis revealed no signficant differences for the US-based regressions (Figure \@ref(fig:hetero-us-plots), SI). Therefore, our study provides no evidence of cross-country differences in the observed associations.

```{r hetero-us, fig.asp = 1, fig.width = 9, include = FALSE, eval = FALSE}
library(patchwork)
names <- c("interd", "conv", "age", "women", "atyp")
plots <- lapply(names, \(x) {
    coeffs %>%
        filter(
            grepl("team", term), 
            grepl(x, model), 
            !grepl("_stem", model)
        ) %>%
        mutate(
            depvar = gsub("fit_([^_]+)(_.*)?", "\\1", model),
            us = ifelse(grepl("us", model), "US", "All"),
        ) %>%
        ggplot(
            aes(
                x = as.numeric(year),
                y = estimate,
                color = us,
            )
        ) +
        geom_hline(
            aes(yintercept = 0, linetype = "Man alone (M)"),
            color = "red"
        ) +
        scale_linetype_manual(values = "dashed") +
        facet_grid(depvar ~ term, labeller = labeller(term = term_labels)) +
        geom_smooth(
            method = "gam",
            formula = y ~ s(x, bs = "cs"),
            aes(weight = 1 / std.error^2),
        ) + 
        theme(axis.title = element_blank())
})
plots_us <- plots[[1]] / plots[[2]] / plots[[3]] / plots[[4]] / plots[[5]] + 
    plot_annotation(tag_levels = "A") +
    plot_layout(guides = "collect")
```


Finally, we have assessed the role of instructor seniority within mixed-gender teams, as reflected in the listing order of instructors in the syllabi. Specifically, we allowed the mixed-gender team coefficients to vary depending on whether the first instructor listed was male ("male leading") or female ("female leading"). The analysis revealed no significant differences in the size or direction of the coefficients for all the selected variables, except for the share of female authors cited. Mixed-gender teams with a female leading instructor tend to cite a higher share of female authors than those with a male leading instructor (Figure \@ref(fig:leading), SI). Therefore, allowing the associations to vary by instructor seniority does not substantially change the results.

## Robustness

The robustness checks conducted, as detailed in the Supplementary Information (Section \@ref(robustness)), show that the main results are consistent across multiple alternative analyses. These included mixed-effects regressions on the full dataset, log-transformations of key dependent variables, quasi-Poisson models for counts of female authors, trimming outliers, and testing different specifications related to instructor composition and seniority.^[We also explored richer specifications that allow for either field-varying or country-varying course level effects. The yields nearly identical results.] Across all these variations, the findings remained broadly stable, demonstrating that the results are not sensitive to specific modeling choices or data subsets.