# Results

## Team formation

We first examined trends in the gender composition of university courses between 2000 and 2019 (Figure \@ref(fig:trends)). In single-instructor courses, the share of female-only courses increased from approximately 30% to 45% between 2000 and 2019. Likewise the share of female two-instructor courses increased from 20% to 35% over the same period. This aligns with the increase in women holding academic positions worldwide.[^women-academia] The share of mixed-gender courses, however, remained relatively stable at 20%, despite the concomitant growing presence of women in academia.

[^women-academia]: In 2000, women comprised approximately 35% of the global academic staff. By 2022, this figure had risen to about 44%, according to data from the World Bank. The data are available at: https://data.worldbank.org/indicator/SE.TER.TCHR.FE.ZS

```{r trends, fig.cap = cap}
cap <- "Trends in the gender composition of university courses, 2000-2019. Left, percentage of solo-taught courses by instructor gender. Right, percentage of two-instructor courses by gender composition. The proportion of mixed-gender teams remains relatively stable. N = 5.1 million courses." # nolint

knitr::include_graphics(
    here::here(
        "output/01_plot_montecarlo_analysis/01_annual_trends.pdf"
    )
)
```


We next compared the observed team composition with a counterfactual in which instructors were randomly matched within institution, academic year, team size, and academic field, while preserving the observed number of teams in each stratum (Sec. \@ref(sec:methods)). This analysis reveals that mixed-gender courses are consistently *underrepresented* in the actual data compared to this gender-neutral matching process, while same-gender courses are consistently *over-represented* (Figure \@ref(fig:montecarlo)). Among two-instructor courses, mixed-gender teams account for 11.4% (male-female) and 9.3% (female-male) of courses, against 19.9% expected under neutral matching, shortfalls of 8.5 and 10.6 percentage points respectively. The corresponding "surplus" falls on same-gender teams, which exceed their simulated share by 9.1 (male-male) and 10.1 (female-female) percentage points. Furthermore, while women's presence grows over time, the gap between observed and expected mixed-gender teams remains remarkably stable after 2005, which is consistent with either persistent barriers or gender-based preferences that may discourage mixed-gender collaboration.

```{r montecarlo, fig.cap = cap}
cap <- "Comparison of gender composition between actual courses and courses simulated with Monte Carlo. Simulations ensure gender neutrality while keeping constant institutional constraints (i.e., number of teams by size, institution, academic year, and 69 academic fields). Simulated mixed-gender teams consistently exceed observed proportions, highlighting persistent barriers to forming gender-diverse teaching teams." # nolint

knitr::include_graphics(
    here::here(
        "output/01_plot_montecarlo_analysis/02_obs_vs_simul.pdf"
    )
)
```

Figure \@ref(fig:montecarlo-by-field) disaggregates the above comparison by academic field, showing that the underrepresentation of mixed teams is consistent across fields, although its size varies substantially. At first glance, the *absolute* shortfall (the gap in percentage points between the simulated and observed share of mixed-gender teams) appears larger in female-rich fields, for example Library Science (57% female instructors; 28 pp shortfall), Medicine (58%; 21 pp), and Education (64%; 19 pp), and is weakly positively correlated with a field's female share across fields ($r = 0.18$). This pattern, however, is partly mechanical: fields with a very skewed gender mix, whether toward women or men, can form few mixed teams in the first place, so there is simply less room for a shortfall. Normalising for this, the *relative* shortfall (the fraction of the expected mixing that fails to materialise) is instead strongly associated with how gender-balanced a field is ($r = 0.47$). In other words, fields closer to a 50/50 instructor mix, such as Media and Communications, English Literature, and Sociology, leave the largest fraction of their potential mixed teams unformed (around 45–60%). This is more consistent with a same-gender preference (homophily) that operates similarly in women-rich and men-rich fields, rather than with women-rich fields behaving differently from men-rich ones.

```{r montecarlo-by-field, fig.cap = cap}
cap <- "This figure disaggregates the comparison of proportions between actual (green) and simulated (orange) mixed-gender (MF/FM) courses by academic field. The shaded bar indicates a 95% confidence level of the proportion." # nolint

knitr::include_graphics(
    here::here(
        "output/01_plot_montecarlo_analysis/03_by_field.pdf"
    )
)
```

Figure \@ref(fig:montecarlo-by-cntry) further shows that the observed proportion of mixed-gender teams consistently fall short of what is expected in the simulations across the US, EU member states, Canada, Great Britain, and the other countries in our sample. Despite temporal variations in the magnitude of these differences, these patterns underscore the robustness of the findings across diverse geographic contexts.

```{r montecarlo-by-cntry, fig.cap = cap, echo = FALSE}
cap <- "This figure disaggregates the comparison of proportions between actual (green) and simulated (orange) mixed-gender courses by country. Simulated mixed-gender teams consistently exceed observed proportions in all countries. EU countries are Austria, Denmark, France, Germany, Ireland, Italy, Netherlands, Poland, Portugal, Spain and Sweden." # nolint

knitr::include_graphics(
    here::here(
        "output/01_plot_montecarlo_analysis/04_by_country.pdf"
    )
)
```


### Pre- and Post- COVID-19 in Europe

Next, we examined whether the composition of teaching teams changed before and after the COVID-19 pandemic, using a recent dataset of 2.2 million European syllabi from 2017–2022 spanning thirty-three European countries. Restricting attention to one- and two-instructor courses with known instructor gender (1.5 million courses), we estimate binomial regressions of team composition on a post-pandemic indicator (2020 onwards), with institution random effects (Sec. \@ref(sec:methods)).

Across countries and academic fields, the total number of courses declined sharply in the post-COVID period (Figure \@ref(fig:covid-trend), in SI). The decline, however, was far from uniform across types of teaching arrangement (Figure \@ref(fig:covid-coeffs)). The clearest shift is in team size: the odds that a course was taught by a single instructor rather than a team fell by about 27% after the onset of the pandemic. Within multi-instructor teams, mixed-gender collaborations also became relatively less common, with the odds of a team being mixed-gender rather than same-gender falling by about 7%. By contrast, the relative share of women among instructors was essentially unchanged (a difference of under 1%, not statistically distinguishable from zero). Taken together, these results indicate that collaborative teaching held up better than solo teaching during the pandemic, but that the collaborations that did form were increasingly same-gender; the overall gender balance of instructors, meanwhile, was largely unaffected.

```{r covid-coeffs, fig.cap = cap}
cap <- paste(
    "**Team composition post-Covid-19 change**. Coefficients and 95% confidence intervals for the post-covid period obtained from the following binomial regressions: (a) probability of single teams being taught by a woman vs. man; (b) probability of courses being taught by one instructor vs. a team (teaching collaboration); (c) probability of teaching teams being mixed-gender vs. same-gender collaborations. The horizontal axis reports the change on the odds scale ($100 \\times (e^{\\beta} - 1)$). All regressions are based on 2.2M syllabi from institutions in thirty-three European countries (excluding teams with three or more collaborators, and authors of unknown gender)." # nolint
)

knitr::include_graphics(
    here::here(
        "output", "02_plot_covid_analysis", 
        "01_covid_coeffs.pdf"
    )
)
```

Figure \@ref(fig:covid-simulations) places these changes against the gender-neutral Monte Carlo benchmark for the same dataset. Consistent with the global analysis, mixed-gender teams are persistently rarer than the simulation predicts throughout the period. The gap between observed and simulated mixing is stable before 2020 but widens in 2022, as the observed share of mixed-gender teams falls further below the simulated share. This pattern is consistent with the pandemic, and the lockdown measures implemented across Europe in particular, having hindered the formation of mixed-gender teaching teams, over and above the mechanical effects of the overall drop in course offerings. We read this as suggestive rather than conclusive, given that it rests on a single post-pandemic year and cannot separate the effect of the pandemic from other concurrent trends.

```{r covid-simulations, fig.cap = cap, fig.width = 7, fig.asp = .45}
cap <- paste(
    "**Post-Pandemic Team Formation**. Simulated gender-neutral vs observed teaching teams for 2.2M syllabi from institutions in thirty-three European countries. Covid-19 period highlighted (2020-2021). This shows that the gap between mixed-gender and same-gender collaborations widens after Covid." # nolint
)
knitr::include_graphics(
    here::here(
        "output/02_plot_covid_analysis/01_covid_sims.pdf"
    )
)
```

## Associations with Course Materials

Figure \@ref(fig:associations) summarises the associations between teaching team composition with five characteristics of course materials: interdisciplinarity, readings age, the share of female authors among cited works, conventionality and atypicality of the readings. Each panel reports estimates along with 95% confidence intervals (CI) from separate linear mixed-effects regressions across all academic years, accounting for unobserved heterogeneity across academic fields and institutions, temporal variation, and other observable course characteristics.

```{r associations, fig.cap = cap}
cap <- paste(
    "**Associations between teaching-team composition and course materials.**",
    "Estimated differences, with 95% confidence intervals, in five characteristics of course materials",
    "(conventionality, share of women authors cited, atypicality, age of readings, and interdisciplinarity)",
    "relative to courses taught by a single male instructor, for single-female, female–female, male–male, and mixed-gender courses.",
    "Each outcome except the share of women authors is expressed as a within-year percentile rank;",
    "the share of women authors is reported in percentage points.",
    "Estimates pool separate linear mixed-effects regressions fitted in each academic year (2000–2019) by inverse-variance weighting,",
    "controlling for country, course level, academic field (random intercept), institution (random intercept),",
    "the probability that the document is a syllabus, STEM status, and the number of readings." # nolint
)
knitr::include_graphics(
    here::here(
        "output/04_make_plots_v2/main_forest_plot.pdf"
    )
)
```

Overall, the results point to a consistent assocation between the gender composition of teaching teams and several dimensions of course materials. First, we observed that the presence of women in teaching teams is associated with the course's interdisciplinarity ranking within the academic year. Relative to single-male courses, the within-year percentile rank of the interdisciplinarity score was highest for mixed-gender courses (1.2 pp, [1.0, 1.5]), followed by female–female courses (1.1 pp, [0.9, 1.3]) and single-female courses (0.2 pp, [0.1, 0.2]). In contrast, the difference for male–male courses was small and statistically uncertain (0.1 pp, [−0.1, 0.2]). 

Second, the presence of women in teaching teams is associated with greater gender represenation of women authors in course readings. Relative to courses taught by one single male instructor, the fraction of women authors cited in readings was 7.6 pp [7.3, 7.8] higher in female–female courses and 7.0 pp [6.9, 7.1] higher in single-female courses. It was 3.0 percentage point higher in mixed-gender courses [2.7, 3.2] and it was lowest in male–male courses (−0.3 pp relative to man-alone, [−0.5, −0.2]).

The analysis of bibliometric measures of readings' novelty and recency yielded a more complex pattern. The presence of male instructors is associated with readings that are both more conventional and more atypical. In contrast, the presence of women instructors is associated with more recent readings. Specifically, Uzzi’s conventionality, expressed as a within-year percentile rank, was higher in male–male courses (0.9 pp, [0.7, 1.0]) and mixed-gender courses (0.9 pp, [0.7, 1.1]), relative to single-male courses, while it was lower in single-female (−0.1 pp, [−0.2, −0.1]) and female–female courses (−0.4 pp, [−0.6, −0.3]). Atypicality, likewise measured as a within-year percentile rank, was lower across all course types, with the largest difference observed for mixed-gender courses (−1.6 pp, [−2.1, −1.0]), followed by female–female (−0.6 pp, [−1.1, −0.1]), male–male (−0.5, [−0.9, −0.1]), and single-female courses (−0.4, [−0.6, −0.2]). Finally, the age of readings, also expressed as a within-year percentile rank, was lower in single-female (−1.3 pp, [−1.4, −1.3]), female–female (−1.6 pp, [−1.8, −1.4]), and mixed courses (−0.5 pp, [−0.7, −0.3]), while there was no clear difference for male–male courses.

```{r temporal-stability, fig.cap = cap}
cap <- paste(
  "**Temporal stability of associations between teaching-team composition and course materials.**",
  "Estimates and 95% confidence intervals for differences relative to courses taught by a single male instructor.",
  "Points show estimates from separate mixed-effects regressions for each academic year; smoothed curves show local estimates averaged across years and weighted by the precision of the yearly estimates."
)

knitr::include_graphics(
    here::here(
        "output/04_make_plots_v2/supp_temporal_stability.pdf"
    )
)
```

To examine whether these associations are stable over time, Figure \@ref(fig:temporal-stability) reports the corresponding year-specific estimates and their smoothed trends. The figure shows that the main differences are generally persistent across the period, while also revealing some temporal variation in their magnitude. In particular, the associations with interdisciplinarity and the representation of women among cited authors show relatively clear changes over time, whereas the differences in conventionality and atypicality are more stable. The temporal patterns are discussed in greater detail alongside the corresponding outcomes below.

## Heterogeneity

Figure \@ref(fig:stem) highlights several key differences between STEM and non-STEM fields, based on separate models estimated for STEM fields compared against results from the full sample.^[
    We converted CIP fields into ISCED-F broad categories and considered STEM the following: (1) Natural Sciences, Mathematics, and Statistics,
    (2) Information and Communication Technologies (ICTs), and (3) Engineering, Manufacturing, and Construction.
]. The gap in female-author citations associated with women instructors remains consistent but is significantly smaller in STEM fields, roughly half the size observed in all fields. The positive association between mixed-gender teams and conventional readings is larger in STEM fields. Conversely, the gap in atypicality widens for women teaching alone, although the effect has been diminishing over time. In contrast, the gap in the age of assigned readings between women and men teaching alone is smaller in STEM. With respect to interdisciplinarity, the estimated differences are small and not consistently significant across years. As a result, the evidence for heterogeneity in interdisciplinarity within STEM is inconclusive.

```{r stem, fig.cap = cap, fig.asp = 1, fig.width = 9}
cap <- paste(
    "**Heterogeneity within STEM Fields**. Estimates and 95% confidence intervals for differences in (1) interdisciplinarity (2) readings' conventionality (3) age of readings (4) female-authored ratio and (5) readings' atypicality. Estimates are from separate mixed-effects regressions for STEM fields vs the full sample, in each academic year, controlling for country, field, institution, course level, probability that the document is a syllabus, and number of readings, with weights proportional to the square root of the number of readings. Smoothed curves show local estimates averaged across years, weighted by the precision (standard errors) of the estimates." #nolint
)

knitr::include_graphics(
    here::here(
        "output/09_heterogeneity_analysis/supp_heterogeneity_stem.pdf"
    )
)
```

We have also explored heterogeneity across countries by fitting separate models for the US and comparing the results against the estimates from the full sample. The US estimates closely track the full-sample estimates for every outcome: the female-author gap for women instructors (4.7 percentage points), the lower age of readings for female-only and female–female courses, the higher conventionality of male–male and mixed-gender courses, and the greater interdisciplinarity of courses involving women are all reproduced within overlapping confidence intervals (Figure \@ref(fig:hetero-us-plots), SI). Therefore, our study provides no evidence of cross-country differences in the observed associations.

Finally, we have assessed the role of instructor seniority within mixed-gender teams, as reflected in the listing order of instructors in the syllabi. Specifically, we allowed the mixed-gender team coefficients to vary depending on whether the first instructor listed was male ("male leading", MF) or female ("female leading", FM). The clearest difference is in the share of female authors cited: mixed-gender teams with a female leading instructor cite about 1.6 percentage points more women authors than those with a male leading instructor [1.3, 1.8]. The remaining outcomes show only small differences by listing order; female-leading teams assign marginally more recent readings and are slightly more interdisciplinary, but these differences are an order of magnitude smaller than the female-author gap (Figure \@ref(fig:leading), SI). Overall, allowing the associations to vary by instructor seniority does not substantially change the results.

### Age and seniority composition

A reviewer raised the concern that the gender gaps could be an artifact of the changing age and seniority composition of the academic workforce rather than team gender composition itself: if senior staff (disproportionately men) teach broad introductory courses while junior women teach specialist courses, the gaps we observe would operate entirely through course level. We cannot observe instructor age directly, but we test the observable implications of this mechanism.

First, if the gaps reflected a seniority-via-course-type sorting, they should collapse once we condition on course level. They do not. Allowing the team contrasts to vary by course level (basic, advanced, graduate) leaves the within-level estimates close to the level-adjusted pooled estimates for the key contrasts; the female-author gap, the lower age of readings for courses involving women, and the mixed-gender conventionality gap all persist within each level rather than shrinking toward zero (Figure \@ref(fig:age-level), SI). The descriptive composition is also inconsistent with the mechanism: the female share of solo-taught courses is essentially flat across levels (43.9% basic, 41.1% advanced, 39.9% graduate), rather than showing the strong concentration of women in specialist courses the account requires.

Second, if the gaps were driven purely by the rising female share of instructors, the level-adjusted gaps should move mechanically with that share over time. Instead, the year-by-year gaps are largely stable and, where they do move, they do not consistently track the female share (which rose from 32.5% in 2000 to 44.2% in 2019). For instance, the female-author gap is, if anything, negatively correlated with the female share, and the interdisciplinarity gap shows a weak or wrong-signed relationship (Figure \@ref(fig:age-trend), SI). Taken together, neither implication of the age/seniority account is borne out, which supports interpreting the associations as reflecting team gender composition rather than a cohort artifact.

## Robustness

The robustness checks conducted, as detailed in the Supplementary Information (Section \@ref(robustness)), show that the main results are consistent across multiple alternative analyses. These included mixed-effects regressions on the full dataset, log-transformations of key dependent variables, quasi-Poisson models for counts of female authors, trimming outliers, and testing different specifications related to instructor composition and seniority.^[We also explored richer specifications that allow for either field-varying or country-varying course level effects. These yield nearly identical results.] Across all these variations, the findings remained broadly stable, demonstrating that the results are not sensitive to specific modeling choices or data subsets.