---
title: "Drivers of mixed-team under-formation across academic fields"
subtitle: "Observed vs. gender-neutral Monte Carlo, by field"
author: "Exploratory notebook"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    df_print: paged
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 7, fig.height = 4.5, dpi = 110
)
```

## Question

Mixed-gender teaching teams occur far less often than a gender-neutral
matching would predict. This notebook asks **what, at the field level,
drives the size of that shortfall**, and in particular:

> Are fields rich with women *more* likely to form mixed-gender teams, or
> *less*?

We compare the **observed** share of mixed-gender two-instructor courses
in each academic field with the share produced by the project's
**gender-neutral Monte Carlo simulation** (`team_rand`), which reshuffles
instructors while holding institutional constraints fixed (team sizes,
institution, year, field). The gap between the two is the field's
"mixed-team shortfall". We then relate that shortfall to each field's
female share of instructors and to its gender *balance*.

```{r libs}
library(dplyr)
library(tidyr)
library(ggplot2)

source(here::here("R/utils.R")) # log_msg(), sep()
source(here::here("R/theme.R")) # theme_custom()
theme_set(theme_custom())
```

## Data

`team_ordered` is the observed team composition; `team_rand` is the
gender-neutral simulated counterpart. Both use the same categories
(`m`, `f`, `mm`, `ff`, `mf`, `fm`), so observed and simulated shares are
directly comparable within a field.

```{r load}
d <- readRDS(here::here("data/processed/syllabi_merged.rds")) |>
  mutate(team_size = nchar(as.character(team_ordered)))

mixed_codes <- c("fm", "mf")

cat("Rows:", format(nrow(d), big.mark = ","), "\n")
cat("Fields:", n_distinct(d$field), " | ISCED groups:", n_distinct(d$isced), "\n")
```

## Field-level measures

**(1) Female share of instructors.** We count every instructor slot in
the observed teams (so a two-woman team contributes two female slots) and
take the share that are women. This is the field's "how rich in women" axis.

```{r female-share}
female_share <- d |>
  transmute(field, isced, slot = strsplit(as.character(team_ordered), "")) |>
  unnest(slot) |>
  group_by(field) |>
  summarise(
    female_share = mean(slot == "f"),
    n_slots      = n(),
    .groups      = "drop"
  )
```

**(2) Observed vs. simulated mixed-team share**, among two-instructor
courses only (mixed teams require exactly two instructors of different
gender). The *shortfall* is `simulated - observed`: how much mixing the
neutral benchmark expects but that we do not see.

```{r mixed-share}
mixed <- d |>
  filter(team_size == 2) |>
  group_by(field, isced) |>
  summarise(
    n2        = n(),
    obs_mixed = mean(team_ordered %in% mixed_codes),
    sim_mixed = mean(team_rand    %in% mixed_codes),
    .groups   = "drop"
  ) |>
  mutate(
    shortfall     = sim_mixed - obs_mixed,          # absolute pp gap
    rel_shortfall = (sim_mixed - obs_mixed) / sim_mixed # fraction of expected mixing lost
  )
```

Combine, keep fields with enough two-instructor courses to be stable, and
add a **balance** index that is 1 for a perfectly 50/50 field and 0 for a
single-sex field.

```{r combine}
res <- mixed |>
  left_join(female_share, by = "field") |>
  filter(n2 >= 200) |>
  mutate(balance = 1 - abs(female_share - 0.5) / 0.5)

cat("Fields analysed:", nrow(res), "\n")
cat("Mean observed mixed share:  ", sprintf("%.1f%%", 100 * mean(res$obs_mixed)), "\n")
cat("Mean simulated mixed share: ", sprintf("%.1f%%", 100 * mean(res$sim_mixed)), "\n")
cat("Mean relative shortfall:    ", sprintf("%.0f%%", 100 * mean(res$rel_shortfall)), "\n")
```

Across fields, roughly **a third of the mixing the neutral benchmark
predicts never materialises** -- the familiar under-formation of
mixed-gender teams, now resolved by field.

## Does female-richness predict the shortfall?

### Naive view: absolute shortfall vs. female share

```{r plot-naive}
ggplot(res, aes(female_share, shortfall)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  geom_point(aes(size = n2), alpha = 0.6, color = "#1b9e77") +
  geom_smooth(method = "lm", formula = y ~ x, color = "#d95f02", fill = "#d95f02", alpha = 0.15) +
  scale_x_continuous(labels = \(x) paste0(100 * x, "%")) +
  scale_y_continuous(labels = \(x) paste0(100 * x, " pp")) +
  scale_size_area(max_size = 6, guide = "none") +
  labs(
    x = "Female share of instructors in the field",
    y = "Mixed-team shortfall (simulated \u2212 observed)",
    title = "Absolute shortfall vs. female share",
    subtitle = "Each point is a field; point size \u221d number of two-instructor courses"
  )
```

```{r cor-naive}
cat("cor(absolute shortfall, female_share) =",
    round(cor(res$shortfall, res$female_share), 3), "\n")
```

The absolute shortfall is **larger in female-rich fields**. But this is
partly mechanical: fields with a very lopsided gender mix (whether toward
women or men) can form few mixed teams in the first place, so there is
less room for a shortfall. The cleaner question uses the *relative*
shortfall and separates "richness in women" from "gender balance".

### Clarified view: relative shortfall vs. female share and balance

```{r plot-balance}
res_long <- res |>
  select(field, isced, rel_shortfall, female_share, balance) |>
  pivot_longer(c(female_share, balance), names_to = "driver", values_to = "x") |>
  mutate(driver = recode(driver,
    female_share = "Female share of instructors",
    balance      = "Gender balance (1 = 50/50)"
  ))

ggplot(res_long, aes(x, rel_shortfall)) +
  geom_point(alpha = 0.55, color = "#1b9e77") +
  geom_smooth(method = "lm", formula = y ~ x, color = "#d95f02", fill = "#d95f02", alpha = 0.15) +
  facet_wrap(~driver, scales = "free_x") +
  scale_x_continuous(labels = \(x) round(x, 2)) +
  scale_y_continuous(labels = \(x) paste0(100 * x, "%")) +
  labs(
    x = NULL,
    y = "Relative shortfall of mixed teams",
    title = "What tracks the relative under-formation of mixed teams?",
    subtitle = "Fraction of the neutral-benchmark mixing that is not realised"
  )
```

```{r cor-clarified}
cat("cor(relative shortfall, female_share) =",
    round(cor(res$rel_shortfall, res$female_share), 3), "\n")
cat("cor(relative shortfall, balance)      =",
    round(cor(res$rel_shortfall, res$balance), 3), "\n")
```

Balance is the strong predictor; the simple female-share correlation is
weak and mostly a by-product of the fact that female-rich fields in this
sample tend to be more lopsided. A regression that includes both makes
this explicit.

```{r regression}
fit <- lm(rel_shortfall ~ female_share + balance, data = res)
summary(fit)
```

`balance` carries a large, significant positive coefficient, while
`female_share` is near zero and not significant once balance is held
constant. In other words, **the more gender-balanced a field is, the
larger the share of potential mixed teams that fails to form** -- being
rich in women does not, by itself, make a field more or less likely to
mix.

## Which fields drive each tail?

```{r tails}
res |>
  arrange(desc(rel_shortfall)) |>
  transmute(
    field, isced,
    female_share  = round(female_share, 2),
    balance       = round(balance, 2),
    obs_mixed     = round(obs_mixed, 3),
    sim_mixed     = round(sim_mixed, 3),
    rel_shortfall = round(rel_shortfall, 2)
  ) |>
  head(10)
```

```{r tails-low}
res |>
  arrange(rel_shortfall) |>
  transmute(
    field, isced,
    female_share  = round(female_share, 2),
    balance       = round(balance, 2),
    obs_mixed     = round(obs_mixed, 3),
    sim_mixed     = round(sim_mixed, 3),
    rel_shortfall = round(rel_shortfall, 2)
  ) |>
  head(10)
```

## Takeaway

- Averaged over fields, about a third of the mixed-gender teaming that a
  gender-neutral benchmark predicts does not occur.
- The **absolute** shortfall correlates *positively* with female share
  (cor = 0.18), but largely because balanced fields have more room to mix
  in the first place; it is partly a mechanical artifact of composition.
- The **honest driver is gender balance, not female-richness**. Fields
  closer to a 50/50 instructor mix leave the largest fraction of their
  potential mixed teams unformed, consistent with same-gender preference
  (homophily) rather than with women-rich fields behaving differently from
  men-rich ones.
- So the answer to "are fields rich with women more likely to form mixed
  teams?" is: **not materially**. What predicts the under-formation of
  mixed teams is how balanced the field is, with female share adding no
  independent signal once balance is accounted for.
